---
title: "BeerAdvocate Data Exploration and Cleaning"
author: "Jakob Johnson"
date: "11/16/2018"
output:
# md_document
html_document:
css: styles.css
---

# BeerAdvocate Data Exploration and Cleaning
For CS 5890 Final Project

# Exploration
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
library(data.table)
library(ggplot2)
library(dplyr)
```

## Import Data

In this exploration we will be using `ggplot2`, `dplyr`, and `data.table` to read in the data.

```{r import1, warning=FALSE}
t1 <- fread("data/beer_reviews-1.csv")
t2 <- fread("data/beer_reviews-2.csv")
t3 <- fread("data/beer_reviews-3.csv")
t4 <- fread("data/beer_reviews-4.csv")
  
ba_raw <- rbind(t1, t2, t3, t4) %>% as_tibble() %>% select(-V1)# import the data as a dplyr tibble
rm(t1, t2, t3, t4)
colnames(ba_raw) # the variables:
```

Let's get rid of the profile name since it won't be useful. 

```{r import2}
ba <- select(ba_raw, -c(review_profilename)) %>% 
  mutate(review_time = review_time)
knitr::kable(head(ba))
```

These data represent `r length(ba$review_overall)` different reviews of `r length(unique(ba$beer_name))` beers from `r length(unique(ba$brewery_name))` different breweries. 

## Plots

```{r plots1, fig.align="center", warning=FALSE}
p1 <- ggplot(ba) +
  stat_bin(aes(x = review_overall), binwidth = 0.25, fill = "#d1730c")
p1
```

```{r plots2, fig.align="center"}

overTime <- tibble(time = c(8, 9, 10, 11, 12,
                            8, 9, 10, 11, 12,
                            8, 9, 10, 11, 12,
                            8, 9, 10, 11, 12,
                            8, 9, 10, 11, 12,
                            8, 9, 10, 11, 12), 
                   ave_rating = c(3.9, 3.8, 3.6, 4.7, 3.4, 
                                  4.0, 3.7, 3.5, 3.6, 3.3, 
                                  4.0, 3.7, 3.5, 3.6, 3.3,
                                  3.5, 3.6, 3.4, 3.5, 3.5, 
                                  3.8, 3.9, 3.8, 3.7, 4.0, 
                                  3.8, 3.9, 3.8, 3.7, 4.0),
                   person = c(1,1,1,1,1,
                              2,2,2,2,2,
                              3,3,3,3,3,
                              4,4,4,4,4,
                              5,5,5,5,5,
                              6,6,6,6,6),
                   abv = c('high','high','high','high','high',
                           'high','high','high','high','high',
                           'high','high','high','high','high',
                           'low','low','low','low','low',
                           'low','low','low','low','low',
                           'low','low','low','low','low'))

time_plot <- ggplot(overTime, aes(x = time, y = ave_rating, color = abv)) + 
  geom_smooth() + labs(title = "Average Rating over a night") + theme_bw()
time_plot

ggsave("process_img/time_vis.jpg", time_plot, width = 6, height = 4, units = 'in')

```

<br>

# Data Cleaning and Grouping

## By Beer Grouping
```{r byBeerGrouping, cache = TRUE}
by_beer <- ba %>%
  group_by(beer_name) %>%
  summarize(beer_id = beer_beerid[1],
            brewery_name = brewery_name[1],
            brewery_id = brewery_id[1],
            beer_style = beer_style[1],
            beer_abv = beer_abv[1],
            mean_overall = mean(review_overall),
            mean_aroma = mean(review_aroma),
            mean_appearance = mean(review_appearance),
            mean_palate = mean(review_palate),
            mean_taste = mean(review_taste),
            n_reviews = length(beer_name))%>%
  filter(n_reviews >= 5)

included_beers <- by_beer$beer_name

ba <- ba %>% filter(beer_name %in% included_beers)

knitr::kable(filter(by_beer, brewery_name == "Uinta Brewing Company") %>% head(.,5))
```

```{r tags, fig.align = 'center'}
pickTag <- function(x){
  check = c("IPA", "Belgian", "Dubbel", "Tripel", "Quad", "wit",  "Stout", "Porter", "Pale Ale", "Lager", "Keller", "Pilsner", "Pilsener", "weizen", "Wheat", "Weiss")
  to =    c("IPA", "Belgian", "Belgian", "Belgian", "Belgian", "Belgian","Stout", "Porter", "Pale Ale", "Lager", "Lager", "Pilsner", "Pilsner", "Wheat Beer", "Wheat Beer", "Wheat Beer")
  
  # tag_maps <- tibble(check = check, to = to)
  
  for(i in 1:length(check)){
      if(grepl(check[i], x)){
        return(to[i])
      }
  }
  return("Other")
}

makeTags <- function(style){
  
  tags <- character(length(style))
    
  for(i in 1:length(style)){
    tags[i] = pickTag(style[i])
  }
  
  return(tags)
}

by_beer <- mutate(by_beer, beer_tag = makeTags(beer_style))

write.csv(by_beer, "data/byBeer.csv")

ggplot(by_beer) +
  stat_count(aes(x = beer_tag, fill = beer_tag)) +
  theme_bw() +
  theme(legend.position = "none") 

other_styles <- filter(by_beer, beer_tag == "Other")$beer_style %>% unique()
other_styles

```

After grouping reviews of the same beer, and removing beers with <5 reviews, we are now down to `r nrow(by_beer)` observations and `r ncol(by_beer)` variables. 

## By Brewery Grouping

```{r byBreweryGrouping, cache = TRUE}
by_brewery <- ba %>%
  group_by(brewery_name) %>%
  summarize(brewery_id = brewery_id[1],
            beers = jsonlite::toJSON(unique(beer_name)), 
            beers_id = jsonlite::toJSON(as.character(unique(beer_beerid))), 
            n_beers = length(unique(beer_name)),
            mean_overall = mean(review_overall),
            mean_aroma = mean(review_aroma),
            mean_appearance = mean(review_appearance),
            mean_palate = mean(review_palate),
            mean_taste = mean(review_taste),
            n_reviews = length(brewery_name)) %>%
  filter(n_beers >= 5)

write.csv(by_brewery, "data/byBrewery-noLocations.csv")

included_breweries <- by_brewery$brewery_id

ba <- ba %>% filter(brewery_id %in% included_breweries)

knitr::kable(sample_n(by_brewery, 5))
```

After grouping reviews from beers at the same brewery, and removing breweries with <5 beers, we are now down to `r nrow(by_brewery)` observations and `r ncol(by_brewery)` variables. 

## CSV Split

In this we will split the original `ba` dataset by brewery and write each to a csv. This will allow for easier and faster loading on the webpage. 

```{r csv_split, eval = FALSE}
ba %>% group_by(brewery_name) %>%
  do(write.csv(., sprintf("data/breweries/%s.csv", .$brewery_id[1])))
```
